A Reproducible Data Analysis Workflow with RMarkdown, Git, Make & Docker

### 2020-04-10 manuscript.Rmd

The translation of computational results into a human-readable summary, for example into a technical report, a presentation, or a journal manuscript, is time-consuming and error-prone. Typical errors result from copy\&paste mistakes, erroneous rounding, or missed updates of the manuscript when the computer code has changed. In order to not only create fully reproducible results but also fully reproducible reports, one needs to resort to the Literate Programming paradigm [@knuth1984literate], in which human-readable language and computer code are intermingled to create dynamic documents whose order follows the logic of thought rather then the order of the computer. R Markdown is a simple markup language to create dynamic documents with embedded chunks of R code that can be exported to standard formats such as documents (docx, rtf, epub), presentations (ppt, html) or websites (html) using the `knitr` package [@xie2015; @xie2019]. Several packages extend the functionality of knitr. Of particular note are the `papaja` package [@papaja2018], which offers additional functions to enable American Psychological Association (APA) style document formatting, including a journal-style final typeset format, and the `stargazer` package [@stargazer2018], which provides journal-ready tables and reports of statistical models.  Figure 1 illustrates R Markdown syntax using the `papaja` package and Figure 2 shows the resulting rendered document.

@@

While our approach was designed to scale well with the complexity of a computing-intense project, we realise that this flexibility may not be straightforward to integrate into a researcher's everyday workflow. There are R packages that implement parts of our workflow and, thus, lower the threshold for adoption when the full flexibility provided by our workflow is not needed. The use of R Markdown within a project, tracked with Git can be simplified with the [`workflowR`](https://github.com/jdblischak/workflowr) package [@workflowr]. The [`drake`](https://github.com/ropensci/drake) package [@drake] is directly inspired by Make and takes an R centric approach, making it especially suited for projects only involving R, but can also handle external dependencies. The [liftR ](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr] and the [`holepunch`](https://karthik.github.io/holepunch/) package [@holepunch] automatize the use of Docker. The former is perfectly compatible with the described workflow and we recommend it to users  who are not comfortable with command-line use of Docker. `holepunch` uses [binder](https://mybinder.org) [@binder2018] to move the analysis to the cloud, so that no local installation of Docker is required. `holepunch` is well suited for simple analysis with low computational demands, because binder's  memory and computing time is limited. There are several alternatives to Docker that manage dependencies on R packages. [`packrat`](https://rstudio.github.io/packrat/) [@packrat] is a way to freeze package version via local copies of packages in the project, but it does not gurantee a given base R version or system dependencies. Similar approaches are taken by [`jetpack`](https://github.com/ankane/jetpack) [@jetpack], [`miniCRAN`](https://cran.r-project.org/web/packages/miniCRAN/vignettes/miniCRAN-introduction.html) [@miniCRAN] and [`checkpoint`](https://cran.r-project.org/web/packages/checkpoint/vignettes/checkpoint.html) [@checkpoint]. The package [`reprex`](https://github.com/tidyverse/reprex) [**repr**oducible **ex**ample, @reprex] is also worth noting, but its scope is limited.

>>>>

@@

While our approach was designed to scale well with the complexity of a computing-intense project, we realise that this flexibility may not be straightforward to integrate into a researcher's everyday workflow. There are R packages that implement parts of our workflow and, thus, lower the threshold for adoption when the full flexibility provided by our workflow is not needed. The use of R Markdown within a project, tracked with Git can be simplified with the [`workflowR`](https://github.com/jdblischak/workflowr) package [@workflowr]. The [`drake`](https://github.com/ropensci/drake) package [@drake] is directly inspired by Make and takes an R centric approach, making it especially suited for projects only involving R, but can also handle external dependencies. The [liftR ](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr] and the [`holepunch`](https://karthik.github.io/holepunch/) package [@holepunch] automatize the use of Docker. The former is perfectly compatible with the described workflow and we recommend it to users  who are not comfortable with command-line use of Docker. `holepunch` uses [binder](https://mybinder.org) [@binder2018] to move the analysis to the cloud, so that no local installation of Docker is required. `holepunch` is well suited for simple analysis with low computational demands, because binder's  memory and computing time is limited. There are several alternatives to Docker that manage dependencies on R packages. [`renv`](https://rstudio.github.io/renv/articles/renv.html) [@renv] is a way to freeze package version via local copies of packages in the project, but it does not gurantee a given base R version or system dependencies. Similar approaches are taken by [`jetpack`](https://github.com/ankane/jetpack) [@jetpack], [`miniCRAN`](https://cran.r-project.org/web/packages/miniCRAN/vignettes/miniCRAN-introduction.html) [@miniCRAN] and [`checkpoint`](https://cran.r-project.org/web/packages/checkpoint/vignettes/checkpoint.html) [@checkpoint]. The package [`reprex`](https://github.com/tidyverse/reprex) [**repr**oducible **ex**ample, @reprex] is also worth noting, but its scope is limited.
